CS182: NBA Scheduler

### 2015-12-10 paper.tex
updates, add final summary

	And of course, each team must play 41 home and 41 away games.

@@

@@

	\begin{figure}
	\caption{Example of current state during $MSP$. The number next to the opposing team corresponds to the number of games left to play against the Knicks.
	min\_count(Knicks)=1 and sum\_count(Knicks, Celtics)=2}
	\includegraphics[scale=0.6]{matchups_game_counts}
	\end{figure}

@@

@@

	The rationale here is the season is typically between 160 and 170 days, meaning ideally each game is every other day. So, we specify the domain of possible dates for a game to be a window of dates centered at that ideal date. The heuristics used were as follows:
	\begin{enumerate}
		\item Forward checking was relatively straightforward in this CSP. We remove any True/False possibilities for any pair of teams if one team has reached the limit of home or away games for that matchup. For any date $d$ we assign for some game number $g$, we remove the dates prior to $d$ for any game after $g$ and any date after $d$ for any game before $g$ for any matchup involving one of these teams.
		\item Choosing what matchup to next assign was the most perplexing part of this problem. We initially chose the team with the most assignments up to that point - essentially trying to assign all home/away matchups for a specific team at a time. The difficulty in this CSP is finding the right assignments for matchups of teams that play 3 times. There's an added complexity of having the right combination of 1 home/2 away and 2 home/1 away matchups that keep each team's total home/away games even. What was originally happening was the DFS would assign every matchup for every team until the last few, but it was impossible to find a right assignment for them because too many teams had their 3-game matchups with a given 2-home or 2-aways. There wasn't a good backtracking method implemented once that happened, so the CSP would stagnate through all possible iterations. Given that, we decided to first assign all matchups between opponents that play an even number of times and only after all those were completed do we start on the 3-game matchups. This was the key to solving this CSP.
		\item Once we selected matchup to assign, we checked which values of True or False were possible: that is, which of the two teams could be the home team.  If a team had reached their limit of home games or away games then there is no ordering to be done because there is at most one option. If both teams could potentially be the home team, we chose to first try the team with the fewer number of home games be the potential home team. Then, for any Home/Away assignment, we had our master dates list with potential dates. We ordered these according to the distance to the ``ideal date'',  2*$g$, where $g$ is the game number we had used to first assign the window of possible dates. The resulting list contained possible assignments: alternating True/False assignments with their associated dates ordered by distance to the ``ideal date''.
		\begin{itemize}
			\item As an example suppose for $(Team_1,Team_2,$g$ =10)$, the window of home dates for $Team_1$ are $(17,19,21,23)$ and the window of home dates for $Team_2$ are $(18,20)$. We want to first try $Team_2$ as the home team because to that point it's had less home assignments, so False would take priority to True. The resulting ordering of the domain would be: (20,False), (19,True), (18,False), (21,True), (23,True), (17,True) - ordering of 19/21 and 17/23 could be flipped based Python's sort implementation.
		\end{itemize}
	\end{enumerate}

@@

>>>>

	Each matchup must be split evenly home/away (or as close as possible in the case of 3 game matchups), and of course each team must play 41 home and 41 away games.

@@

	We don't concern ourselves as much with this, we will just randomly generate some set of home dates. More details will follow below.

@@

@@

	\begin{figure}
	\caption{Example of current state during $MSP$. The number next to the opposing team corresponds to the number of games left to play against the Knicks. As Celtics still have to play Knicks 3 times, this state is inconsistent.
	min\_count(Knicks)=1 and sum\_count(Knicks, Celtics)=2}
	\includegraphics[scale=0.6]{matchups_game_counts}
	\end{figure}

@@

	The rationale here is the season is typically between 160 and 170 days, meaning ideally each game is every other day. So, we specify the domain of possible dates for a game to be a window of dates centered at that ideal date. The heuristics used were as follows:
	\begin{enumerate}
		\item Forward checking was relatively straightforward in this CSP. We remove any True/False possibilities for any pair of teams if one team has reached the limit of home or away games for that matchup. For any date $d$ we assign for some game number $g$, we remove the dates prior to $d$ for any game after $g$ and any date after $d$ for any game before $g$ for any matchup involving one of these teams.
		\item Choosing what matchup to next assign was the most perplexing part of this problem. We initially chose the team with the most assignments up to that point - essentially trying to assign all home/away matchups for a specific team at a time. The difficulty in this CSP is finding the right assignments for matchups of teams that play 3 times. There's an added complexity of having the right combination of 1 home/2 away and 2 home/1 away matchups that keep each team's total home/away games even. What was originally happening was the DFS would assign every matchup for every team until the last few, but it was impossible to find a right assignment for them because too many teams had their 3-game matchups with a given 2-home or 2-aways. There wasn't a good backtracking method implemented once that happened, so the CSP would stagnate through all possible iterations. Given that, we decided to first assign all matchups between opponents that play an even number of times and only after all those were completed do we start on the 3-game matchups. This was the key to solving this CSP.
		\item Once we selected the matchup to assign, we checked which values of True or False were possible: that is, which of the two teams could be the home team.  If a team had reached their limit of home games or away games then there is no ordering to be done because there is at most one option. If both teams could potentially be the home team, we chose to first try the team with the fewer number of home games be the potential home team. Then, for any Home/Away assignment, we had our master dates list with potential dates. We ordered these according to the distance to the ``ideal date'',  2*$g$, where $g$ is the game number we had used to first assign the window of possible dates. The resulting list contained possible assignments: alternating True/False assignments with their associated dates ordered by distance to the ``ideal date''.
		\begin{itemize}
			\item As an example suppose for $(Team_1,Team_2,$g$ =10)$, the window of home dates for $Team_1$ are $(17,19,21,23)$ and the window of home dates for $Team_2$ are $(18,20)$. We want to first try $Team_2$ as the home team because to that point it's had less home assignments, so False would take priority to True. The resulting ordering of the domain would be: (20,False), (19,True), (18,False), (21,True), (23,True), (17,True) - ordering of 19/21 and 17/23 could be flipped based on Python's sort implementation.
		\end{itemize}
	\end{enumerate}

@@

    \section{Summary}
    We approached solving this problem in two different ways - using an ILP solver and writing our own CSP solver. We wanted to see if we could beat the ILP solver - but since switching to GUROBI it would seem to be infeasible using pure Python. However, we could reasonably solve the problem in under 5 minutes if a solution for the dates exists. There's certainly many more areas where we could have explored:

    \begin{itemize}
    	\item Finding a good heuristic for the $M$ games in $N$ nights constraint. If we could provide this then it should be possible to avoid getting stuck iterating in a BFS-like manner once the constraint gets broken.
    	\item Extending the model to take into account more realistic scenarios - for example the Lakers and Clippers share a home arena.
    	\item Adding to the cost function some sense of ``primetime'' games, where teams with a better record or larger market playing in those games get a higher score.
    	\item Minimizing long road trips.
    \end{itemize}

    Generally, all of these problems involve playing around with the cost/heuristic function. It was a great experience in terms of using the tools we learned and applying them to a very specific problem - and it shows how knowledge about the problems helps in forming heuristics.
